Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
LUCAS more edits to your section, of course so free to change them back
fixed bug in previous commit

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\endingSystem}{\textsc{DreamCoder}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
% Use the following line for the initial blind version submitted for review:
\usepackage[nohyperref]{icml2018}

@@

This work proposes \emph{learning} the DSL.
We consider the setting where we have a collection of related programming tasks,
each specified by a set of input/output examples.
We do not assume that the tasks are annotated with
ground-truth programs.
Our algorithm, called \system,
does three things:
it finds programs that solve each of the tasks;
it creates its own DSL by discovering and reusing
domain-specific subroutines;
and it learns a neural network that
efficiently searches for programs in the DSL.
The purpose of the DSL
is to expose the commonalities across the programs that solve tasks within a domain,
letting the agent solve other related program learning problems.

@@

How should we perform inference for the model
in~\ref{graphicalModel}(a)?
We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
which alternates between ``waking'', which draws samples from a recognition model (our $q$);
and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
The wake cycle can be thought of active problem solving,
guided by the current generative and recognition models.
A sleep cycle means  updating these models so that the DSL and  recognition model
more closely match the distribution of tasks.
An important difference from the Helmholtz machine
is that our setup has two distinct sleep phases: a phase where we update the generative model
$(\mathcal{D},\theta)$, which we call \textbf{Sleep-G},
and a second phase where we update the recognition model $q$,
which we call \textbf{Sleep-R}.
This wake/sleep style of inference is the reason why we call our approach \endingSystem.

@@

\subsection{String editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string routines provided by Python --
and recover many of the higher-level building blocks that have made these
other system successful.
Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \endingSystem.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide list/string indexing primitives (\code{slice}, \code{nth}),
basic arithmetic, and functions for splitting a string into a list (\code{split})
or turning a list into a string (\code{join}).
We automatically generated  600 string editing tasks along with input/output examples (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
At first, \system cannot find any programs that
work at all for most of the tasks.
After three wake/sleep iterations,
it assembles a DSL (Tbl.~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

@@

  \begin{minipage}{7.5cm}
 $f($\code{s}$) = $   \code{(}$f_3$\code{ (split '>' (upper s)) '<')}
    \end{minipage}
  \caption{Example string edit task (top) and program  \system writes (bottom). $f_3$ is a new DSL subroutine discovered by \system, defined in Tbl.~\ref{textPrimitives}}\label{exampleTextProblem}
  \end{figure}

@@

We found  this domain  difficult to tackle without starting from 
the rich DSL. Our
algorithm, like human learners, requires a spectrum of problems ranging from
easy to hard. Without this curriculum, we cannot get off the ground.
%The behavior with the base DSL illustrates this phenomena  because the initial set of
%primitives is not paired with a sufficiently large and graded curriculum.

@@

 \section{Contribution and Outlook}
 We contributed a model, \system, that learns to program by
 bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.
  Alongside this DSL, we train a bottom-up
 neural recognition model that learns how to efficiently deploy the
 DSL and hone in on the right programs.  We believe that this
 integration of top-down symbolic representation and bottom-up neural
 networks -- both of them learned -- is a useful general approach for
 getting program induction off the ground.

>>>>

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
% Use the following line for the initial blind version submitted for review:
\usepackage[nohyperref]{icml2018}

@@

This work proposes \emph{learning} the DSL.
We consider the setting where we have a collection of related programming tasks,
each specified by a set of input/output examples.
We do not assume that the tasks are annotated with
ground-truth programs.
Our algorithm, called \systemEnding,
does three things:
it finds programs that solve each of the tasks;
it creates its own DSL by discovering and reusing
domain-specific subroutines;
and it learns a neural network that
efficiently searches for programs in the DSL.
The purpose of the DSL
is to expose the commonalities across the programs that solve tasks within a domain,
letting the agent solve other related program learning problems.

@@

How should we perform inference for the model
in~\ref{graphicalModel}(a)?
We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
which alternates between ``waking'', which draws samples from a recognition model (our $q$);
and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
The wake cycle can be thought of active problem solving,
guided by the current generative and recognition models.
A sleep cycle means  updating these models so that the DSL and  recognition model
more closely match the distribution of tasks.
An important difference from the Helmholtz machine
is that our setup has two distinct sleep phases: a phase where we update the generative model
$(\mathcal{D},\theta)$, which we call \textbf{Sleep-G},
and a second phase where we update the recognition model $q$,
which we call \textbf{Sleep-R}.
This wake/sleep style of inference is the reason why we call our approach \systemEnding.

@@

\subsection{String editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string routines provided by Python --
and recover many of the higher-level building blocks that have made these
other system successful.
Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \systemEnding.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide list/string indexing primitives (\code{slice}, \code{nth}),
basic arithmetic, and functions for splitting a string into a list (\code{split})
or turning a list into a string (\code{join}).
We automatically generated  600 string editing tasks along with input/output examples (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
At first, \system cannot find any programs that
work at all for most of the tasks.
After three wake/sleep iterations,
it assembles a DSL (Tbl.~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

@@

  \begin{minipage}{7.5cm}
 $f($\code{s}$) = $   \code{(}$f_3$\code{ (split '>' (upper s)) '<')}
    \end{minipage}
  \caption{Example string edit task (top) and program  \system writes (bottom). $f_3$ is a new DSL subroutine discovered by \systemEnding, defined in Tbl.~\ref{textPrimitives}}\label{exampleTextProblem}
  \end{figure}

@@

We found  this domain  difficult to tackle without starting from 
the rich DSL. Our
algorithm, like human learners, requires a spectrum of problems ranging from
easy to hard.
When starting with the base DSL,
we found that there were not enough steppingstones in the curriculum to
get \system off the ground.

%The behavior with the base DSL illustrates this phenomena  because the initial set of
%primitives is not paired with a sufficiently large and graded curriculum.

@@

 \section{Contribution and Outlook}
 We contributed a model, \systemEnding, that learns to program by
 bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.
  Alongside this DSL, we train a bottom-up
 neural recognition model that learns how to efficiently deploy the
 DSL and hone in on the right programs.  We believe that this
 integration of top-down symbolic representation and bottom-up neural
 networks -- both of them learned -- is a useful general approach for
 getting program induction off the ground.
